Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-03 paper/disc20clustering.tex
Bringing back the old LB

\begin{proof}
	Fix any online algorithm \ALG{}.
	Assume that $k$ is odd.
	We show how to proceed with even $k$ later in this proof.
	Let $I(z)$ denote the cluster where the node $z$ was located initially.
	Throughout the construction, we maintain ground sets for each node.
	Initially all nodes are \emph{alone} (they belong to their own trivial ground sets).

>>>>

\begin{proof}
	\maciek{Better name than $S_1$?}
	Fix any online algorithm \ALG{}.
	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
	\maciek{A problem with the definition! The first pair does not have the same initial location of its nodes.}
	Initially all nodes are single (they belong to their own trivial ground sets).

	We construct a ground set containing this alone node (and some other node).
	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S_1$.
	First,
	we construct a ground set $\{x_0, y_0\}$.
	Since \ALG{} has them split, the adversary produces an external request.
	Since $S_1$ contains a ground set of size $k$, \ALG{} must collocate $\{x_0, y_0\}$ in another cluster.
	To preserve a feasible partition of nodes into clusters,
	\ALG must replace $x_0$ with another single node $x_1$.

	We continue to construct ground sets containing the current alone node from $S_1$.
	Precisely, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t.~$I(C) = I(x_1)$.
	To preserve a feasible partition of nodes into clusters,
	\ALG{} must collocate $x_1$ and $C$ in a~cluster other than $S_1$ and
	consequently replace $x_1$ with another alone node $x_2$.

	We end this procedure prematurely (before all nodes visit $S_1$) to keep \OPT's cost low.
	We repeat merging the component of single node $x_i$ on $S_1$ and the largest component $C'$ s.t.~$I(C')=I(x_i)$,
	until there are only two single nodes left that  originate from the same cluster.
	Formally, we denote these two single nodes by $x^*, y^*$, and we have $I(x^*) = I(y^*)$, and~for any other pair of single nodes
	$x'$ and $y'$,
	we have $I(x') \neq I(y')$.
	Note that at this point there are at most $\ell+1$ single nodes left,
	otherwise, there would be more pairs of single nodes that were initially in the same cluster.

	Finally, we bound the cost of \OPT.
	Given this sequence of requests,
	\OPT migrates $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
	swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
	Hence,
	OPT pays for $2$ node swaps and
	\ALG{} incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
	therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate. Don't use the word "exclude" unless very obvious. Just sum up the real cost.}.
	Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.

	\maciek{BELOW OLD VERSION. TODO: INCORPORATE BETTER SENTENCES}

	Fix any online algorithm \ALG{}.
	Assume that $k$ is odd.
	We show how to proceed with even $k$ later in this proof.
	Let $I(z)$ denote the cluster where the node $z$ was located initially.
	Throughout the construction, we maintain ground sets for each node.
	Initially all nodes are \emph{alone} (they belong to their own trivial ground sets).
